---
title: "Reproducible Research: Peer Assessment 1 "
output: 
  html_document:keep_md: true
----

```{r opts, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "figure/"
)
```


## Loading and preprocessing the data.

```{r}
library(plyr)
library(lattice)
setwd("F:/coursera/repos/RepData_PeerAssessment1/")
padata<-read.csv("activity.csv")
```


## What is mean total number of steps taken per day?

```{r}
###Total number of steps per day
stepsPerDay<-aggregate(steps~date,na.omit(padata),sum)
head(stepsPerDay)
```

```{r, echo=FALSE }
###histogram of the total number of steps taken each day
hist(stepsPerDay$steps,
     main = "Histogram of total steps taken each day - (nulls included)",
     xlab="Steps per day")
```

###Calculate and report the mean and median of the total number of steps taken per day

```{r}
#mean

stepsPerDayMean <- mean(stepsPerDay$steps)
head(stepsPerDayMean)

#median

stepsPerDayMedian <- median(stepsPerDay$steps)
head(stepsPerDayMedian)
```

## What is the average daily activity pattern?

###Plot interval vs Average steps
```{r, echo=FALSE}
intervalStepsMean<-aggregate(steps~interval,data=padata,mean)
plot(intervalStepsMean,type='l',
     col="blue",
     main="Average Number of Steps across all days")
```

###Determine maximum 5 minute interval across all days
```{r}
intervalStepsMean[which.max(intervalStepsMean$steps),1]

```

## Inputing missing values


```{r}
###Total number of missing values in the dataset
sum(is.na(padata))

###New dataset with missing values filled in.
padata1<-padata
padata1$steps[is.na(padata1$steps)] <- intervalStepsMean$steps[is.na(padata1$steps)]
head(padata1)

###Calculate the total number of steps taken per day
stepsPerDay1<-aggregate(steps~date,data=padata1,sum)
head(stepsPerDay1)

```

```{r, echo=FALSE }
#histogram of the total number of steps taken each day
hist(stepsPerDay1$steps,
     main = "Histogram of total steps taken each day - (nulls replaced)",
     xlab="Steps per day")
```

```{r}
###Calculate and report the mean and median of the total number of steps taken per day

####mean
stepsPerDayMean1 <- mean(stepsPerDay1$steps)
stepsPerDayMean1

####median
stepsPerDayMedian1 <- median(stepsPerDay1$steps)
stepsPerDayMean1

## Differences between estimates of filled na with na values

###Difference in Steps per Day Mean
stepsPerDayMean1 - stepsPerDayMean

###Difference in Steps per Day Median
stepsPerDayMedian1 - stepsPerDayMedian

```
##Are there differences in activity patterns between weekdays and weekends?
###Create a new factor variable withlevels - "weekday" and "weekend" 
```{r}
padata1$day<-as.factor(weekdays(as.Date(padata1$date)))
padata1 <- within(padata1, levels(day)[grep("S+",levels(padata1$day))] <- "Weekend")
padata1 <- within(padata1, levels(day)[levels(padata1$day) != "Weekend"] <- "Weekday")
padata1$date<-NULL
```
###Plot time series data - interval vs steps by Weekend, Weekday
```{r , echo = FALSE }
padata1<-aggregate(steps~interval+day,data=padata1,mean)
xyplot(steps ~ interval | day, data=padata1, type = "l", layout = c(1, 2), 
       xlab = "Interval", ylab = "Number of steps")
```

